---
title: "ml Challenge - Kiszkurno Miguel - Exploratory data analysis"
output:
  html_document:
    df_print: paged
---

# Setup

```{r clean up memory}
# memory clean up
rm(list = ls()) # remove all objects
gc() # garbage collection


```

```{r import libraries, warning=FALSE, message=FALSE}

# Import libraries

library(tidyverse)
library(corrr)
library(knitr)
library(kableExtra)
library(GGally)
library(tidymodels)
library(rsample)
library(ggplot2)

```



```{r setup things}   

#Change this if you want to run it
knitr::opts_knit$set(root.dir = "/Users/miguelkiszkurno/Documents/metlife/test-docker/src")

#seed to be used in the future
g_seed = 730 

```


```{r load dataset}

#Load the dataset
ds <- read.table(paste0("/Users/miguelkiszkurno/Documents/metlife/test-docker/datasets/dataset.csv"),
                        sep=",", dec=".", header = TRUE, fill = TRUE)

# fijamos semilla
set.seed(g_seed)

# Split the data
train_test <- initial_split(ds, prop = 0.9)
ds <- training(train_test)
ds_test <- testing(train_test)

# Check dimensions
ds %>%
  dim_desc() 
ds_test %>%
  dim_desc() 

#General view of the dataset
glimpse(ds)

```

# Preliminary analysis


```{r unique and missing values}

#Gather information
exp_table =  ds %>%
                gather(., 
                      key = "variables", 
                      value = "values") %>% # Grouping by variable
                group_by(variables) %>% 
                summarise(unique_values = n_distinct(values),
                percentaje_missing = sum(is.na(values))/nrow(ds)*100) %>% 
                arrange(desc(percentaje_missing), unique_values) # order by percentaje

#View the table
exp_table

```

```{r categorical variables}

# Want to check factor variables
regions <- data.frame(campo = "region", valor = unique(ds$region))
smokers <- data.frame(campo = "smoker", valor = unique(ds$smoker))
sex <- data.frame(campo = "sex", valor = unique(ds$sex))

# Combine the datasets to see everything together
df_unique_values <- do.call(rbind, list(regions, smokers, sex))

# View the dataset
df_unique_values


```


```{r factor variables conversion}

transform_factor_variables <- function (data_set){
  
  data_set$region <- factor(data_set$region)
  data_set$sex <- factor(data_set$sex)
  data_set$smoker <- data_set$smoker == "yes"

  return (data_set)
}

ds <- transform_factor_variables(ds)
ds_test <-transform_factor_variables (ds_test)

# Another quick look to the resulting dataset
glimpse(ds)
```


```{r descriptive metrics}

#Check some standard descriptive metrics of the dataset.
ds_metrics <- ds %>%
      select_if(is.numeric) %>% 
      gather(., 
             key = "variable", 
             value = "values") %>% # agrupamos por las variables del set
             group_by(variable) %>%
                  summarise(min = min(values),
                            mean = mean(values),
                            median = median(values),
                            sd = sd(values),
                            max = max(values) ) %>% 
                                      arrange(variable) # orden
ds_metrics

```


```{r relation between numeric variables (GGPairs), message=FALSE, warning=FALSE, fig.width=8, fig.height=8}

#Miro la relación entre las variables 
numericas <- ds %>%
  select(where(is.numeric), sex)

ggpairs(numericas,  mapping = aes(color = sex))

numericas <- ds %>%
  select(where(is.numeric), smoker)

ggpairs(numericas,  mapping = aes(color = smoker))

numericas <- ds %>%
  select(where(is.numeric), region)
ggpairs(numericas,  mapping = aes(color = region))

```

```{r correlation matrix, message=FALSE}

ds %>% 
 correlate() %>% 
  shave() %>% 
  fashion() 

ds %>% 
 correlate() %>% 
  rplot()

```



$ age      <int> 19, 18, 28, 33, 32, 31, 46, 37, 37, 60, 25, 62, 23, 56, 27, 19, 52, 23, 56, 30, 60, 30, 18, 34, 37, 59, 63, 55, 23, 31, 22, 18, 19, 63, 28, 19, 62, 26, …
$ bmi      <dbl> 27.900, 33.770, 33.000, 22.705, 28.880, 25.740, 33.440, 27.740, 29.830, 25.840, 26.220, 26.290, 34.400, 39.820, 42.130, 24.600, 30.780, 23.845, 40.300, …
$ children <int> 0, 1, 3, 0, 0, 0, 1, 3, 2, 0, 0, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 1, 0, 1, 2, 3, 0, 2, 1, 2, 0, 0, 5, 0, 1, 0, 3, 0, 1, 0, 0, 2, 1, 2, 1, 0, 2, 0, 0, 1, 0,…


```{r target variable boxplots}


# Charges by sex
ggplot(ds, aes(x = sex, y = charges, group = sex, fill = sex))+
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 70000))


# Charges by region
ggplot(ds, aes(x = region, y = charges, group = region, fill = region))+
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 70000)) # definimos escala del eje y


# Charges by region smoker
ggplot(ds, aes(x = smoker, y = charges, group = smoker, fill = smoker))+
  geom_boxplot() +
  scale_y_continuous(limits = c(0, 70000)) # definimos escala del eje y


```




```{r outliers superior threshold}
limite_superior_outliers = IQR(ds$charges) * 1.5 + quantile(ds$charges, 0.75)[[1]]
limite_superior_outliers
```

```{r outliers}
outliers_charges <- ds %>% filter(charges>limite_superior_outliers) %>% arrange(., desc(charges))
outliers_charges
```

```{r outliers inferior threshold}
limite_inferior_charges = quantile(ds$charges, 0.25)[[1]] - IQR(ds$charges) * 1.5
limite_inferior_charges 
```


# Linear Regression: Predict charges using all the features

## Model Fiting
```{r model charges ~ all data}
			
# fit the model
mdl_all = lm(formula = charges ~ ., data = ds)

# View the model summary (dataframe format)
tidy_mdl_all <- tidy(mdl_all, conf.int = TRUE) %>% arrange(p.value)
tidy_mdl_all

# View the model summary (plain text format)
summary_mdl_all = summary(mdl_all)
summary_mdl_all

```

## Model diagnosis

```{r model charges ~ all diagnosis}

# Plot the residuals
plot(mdl_all)

```


# Linear Regression: Predict log(charges) using all the features

## Previous calculations

```{r calculate log.charges}

# Will use the log to normalize the charges field
ds = ds %>% 
  mutate(log.charges = log(charges))

head(ds)

```


```{r view distributions}

#Charges (without applying log)
ggplot(data = ds, aes(x = round(charges))) + 
  geom_histogram(col = "white", aes( fill = ..count..), alpha = 0.75) +
  labs(title = "charges Histogram") +
  labs(x = "charges") +
  theme_bw()

#Charges (applying log)
ggplot(data = ds, aes(x = log.charges)) + 
  geom_histogram(col = "white", aes( fill = ..count..), alpha = 0.75) +
  labs(title = "log.charges Histogram") +
  labs(x = "log.charges") +
  theme_bw()

```

## Model fitting 

```{r Log(charges) ~ all data}

#Fit the model using log.charges as target and all the other features as predictors
mdl_log_all = lm(formula = log.charges ~ age + sex + bmi + children + smoker + region, data = ds)
mdl_log_all

# View the model summary (plain text format)
summary_mdl_log_all = summary(mdl_log_all)
summary_mdl_log_all

# View the model summary (dataframe format)
tidy_summary_mdl_log_all <- tidy(summary_mdl_log_all, conf.int = TRUE) %>% arrange(p.value)
tidy_summary_mdl_log_all

```


```{r plot coeficients significance model charges ~ all}

# Coeficients plot
ggplot(tidy_summary_mdl_log_all, aes(estimate, term, xmin = conf.low, xmax = conf.high, height = 0)) +
  geom_point(color = "forestgreen",size=2) +
  geom_vline(xintercept = 0, lty = 4, color = "black") +
  geom_errorbarh(color = "forestgreen", size=1) +
  theme_bw() +
  labs(y = "Coeficientes β", x = "Estimación")

```

## Model diagnosis

```{r model Log(charges) ~ all data}

# Plot the residuals
plot(mdl_log_all)

```
# Linear Regression: Adding iteration between age and smoker (mdl_log_all_it_smoker_age)
 
## Model Fiting


```{r model log.charges ~ sex + bmi + children + smoker * age + region}

# Fit the model
mdl_log_all_it_smoker_age = lm(formula = log.charges ~ sex + bmi + children + smoker * age + region, data = ds)
mdl_log_all_it_smoker_age

# View the model (dataframe format)
summary_mdl_log_all_it_smoker_age = summary(mdl_log_all_it_smoker_age)
summary_mdl_log_all_it_smoker_age

# View the model (plain text format)
tidy_summary_mdl_log_all_it_smoker_age <- tidy(summary_mdl_log_all_it_smoker_age, conf.int = TRUE) %>% arrange(p.value)
tidy_summary_mdl_log_all_it_smoker_age

```

## Model diagnosis

```{r model model log.charges ~ sex + bmi + children + smoker * age + region diagnosis}

# Plot the residuals
plot(mdl_log_all_it_smoker_age)

```

# Linear Regression: Predict Log(charges) Adding iteration between smoker and bmi (mdl_log_all_it_smoker_age_bmi)


```{r model log.charges ~ sex + smoker * bmi + children + smoker * age + region}

# Fit the model
mdl_log_all_it_smoker_age_bmi = lm(formula = log.charges ~ sex + smoker * bmi + children + smoker * age + region, data = ds)
mdl_log_all_it_smoker_age_bmi

# View the model (plain text format)
summary_mdl_log_all_it_smoker_age_bmi = summary(mdl_log_all_it_smoker_age_bmi)
summary_mdl_log_all_it_smoker_age_bmi

# View the model (dataframe format)
tidy_summary_mdl_log_all_it_smoker_age_bmi <- tidy(summary_mdl_log_all_it_smoker_age_bmi, conf.int = TRUE) %>% arrange(p.value)
tidy_summary_mdl_log_all_it_smoker_age_bmi


```

## Model diagnosis
```{r model log.charges ~ sex + smoker * bmi + children + smoker * age + region diagnosis}
plot(mdl_log_all_it_smoker_age_bmi)

```


# Linear Regression: Predict Log(charges) Adding iteration between age and bmi (mdl_log_all_it_age_bmi)
 

```{r model log.charges ~ sex + age * bmi + children + smoker + region}
mdl_log_all_it_age_bmi = lm(formula = log.charges ~ sex + age * bmi + children + smoker + region, data = ds)
mdl_log_all_it_age_bmi


# View the model summary (plain text format)
summary_mdl_log_all_it_age_bmi = summary(mdl_log_all_it_age_bmi)
summary_mdl_log_all_it_age_bmi

# View the model summary (dataframe format)
tidy_summary_mdl_log_all_it_age_bmi <- tidy(summary_mdl_log_all_it_age_bmi, conf.int = TRUE) %>% arrange(p.value)
tidy_summary_mdl_log_all_it_age_bmi

```

## model diagnosis
```{r model log.charges ~ sex + age * bmi + children + smoker + region diagnosis}

# Plot the residuals
plot(mdl_log_all_it_age_bmi)

```

# Linear Regression: Predict charges Adding iteration between smoker, bmi and Age (mdl_all_it_smoker_age_bmi)


```{r model charges ~ sex + smoker * bmi + children + smoker * age + region}
mdl_all_it_smoker_age_bmi = lm(formula = charges ~ sex + smoker * bmi + children + smoker * age + region, data = ds)
mdl_all_it_smoker_age_bmi

# View the model summary (plain text format)
summary_mdl_all_it_smoker_age_bmi = summary(mdl_all_it_smoker_age_bmi)
summary_mdl_all_it_smoker_age_bmi

# View the model summary (dataframe format)
tidy_summary_mdl_all_it_smoker_age_bmi <- tidy(summary_mdl_all_it_smoker_age_bmi, conf.int = TRUE) %>% arrange(p.value)
tidy_summary_mdl_all_it_smoker_age_bmi


```

## Model diagnosis
```{r model charges ~ sex + smoker * bmi + children + smoker * age + region diagnosis}

# Plot the residuals
plot(mdl_all_it_smoker_age_bmi)

```



# Model Comparison

## Evaluation in training data

```{r model comparison}

get_metrics_from_model <-function(model, test_data = NULL){
  
  target = model[["terms"]][[2]]
  formula = deparse(model[["terms"]][[3]])
  mdl_name = paste0(target, "~", formula)

  if (is.null(test_data)){
    cat("No text data")
    eval <- broom::augment(model, ds)
  }
  else{
    eval <- broom::augment(model, newdata = test_data)
  }

    
  if (target == "charges"){
    metrics = metrics(data = eval, truth = charges, estimate = .fitted) %>% mutate(.estimate = round(.estimate, 4)) %>% mutate(model = mdl_name)
  }
  else{
    # log models require to do the exp
    eval = eval %>%  mutate(fitted_antilog = exp(.fitted))
    
    # Compute RMSE, R2 and MAE for charge (not logcharges) so we can compare all the models
    metrics = metrics(data = eval, truth = charges, estimate = fitted_antilog) %>% mutate(.estimate = round(.estimate, 4)) %>% mutate(model = mdl_name) 
  }
    
  return (metrics)
}

# all the models I want to compare in the models list
models <- list(mdl_all = mdl_all, 
               mdl_log_all = mdl_log_all, 
               mdl_log_all_it_smoker_age = mdl_log_all_it_smoker_age,
               mdl_log_all_it_smoker_age_bmi = mdl_log_all_it_smoker_age_bmi, 
               mdl_log_all_it_age_bmi = mdl_log_all_it_age_bmi, 
               mdl_all_it_smoker_age_bmi = mdl_all_it_smoker_age_bmi
               )
metrics_total <- list()

for (m in models) {
 
  metrics = get_metrics_from_model (m)
  metrics_total <- bind_rows(metrics_total, metrics)
}
#Compare RMSE
metrics_total  %>%
  filter(.metric == "rmse") %>%
  arrange(.estimate)

#Compare MAE
metrics_total  %>%
  filter(.metric == "mae") %>%
  arrange(.estimate)

#Compare R-Squared
metrics_total  %>%
  filter(.metric == "rsq") %>%
  arrange(desc(.estimate))

```

### Para Test

Vamos a calcular el MAE, RMSE y R2 para los 6 modelos

```{r comparo entrenamiento 2}

metrics_total <- list()

for (m in models) {
 
  metrics = get_metrics_from_model (m, ds_test)
  metrics_total <- bind_rows(metrics_total, metrics)
}
#Compare RMSE
metrics_total  %>%
  filter(.metric == "rmse") %>%
  arrange(.estimate)

#Compare MAE
metrics_total  %>%
  filter(.metric == "mae") %>%
  arrange(.estimate)

#Compare R-Squared
metrics_total  %>%
  filter(.metric == "rsq") %>%
  arrange(desc(.estimate))

```
